Where Do Events Come From?

### 2018-09-27 paper.tex

The addition of asynchronous ISRs now poses real threats concerning race
conditions since they interrupt programs at arbitrary points and could possibly
share variables with synchronous code.
%
Listing~\ref{lst.usart} illustrates a minimum serial communication (USART)
application (ln~15--24) that consumes incoming bytes (ln~19--23) as they arrive
(ln~18).
%
The USART API (ln~1--4) exposes an input event to signal incoming data (ln~3)
and a buffer to prevent data loss (ln~4).
The buffer is indispensable because the microcontroller might not cope with the USART 
speed.
%
The (unsafe) driver ISR (ln~6--13) simply appends incoming bytes to the end of the
buffer (ln~9) and signals the application (ln~11).
%
As a possible race condition, note that the ISR may fire and append a new byte
to the buffer (ln~9) just before the application clears it (ln~23), in which
case the new byte will be lost forever.

\CEU treats a \code{spawn~async/isr} as a block of code that runs in parallel with
the rest of the program (hence the prefix \code{spawn}).
%
This way, it is clear for the compiler that the accesses to \code{rx\_buf} in
Listing~\ref{lst.usart} may lead to race conditions (e.g., ln~9,23), and \CEU
raises a compile-time error.
%
The programmer is responsible to protect concurrent memory accesses with
\code{atomic} blocks, which disables interrupts for a short period
of time.

@@

However, we do not expect that application programmers should be required to
resolve race conditions.
\CEU supports \code{code} reactive abstractions (similar to coroutines) that help hiding
concurrency issues inside drivers.
%
Listing~\ref{lst.usart.code} changes the USART API (ln 1--3) to expose a code
abstraction that expects a reference to a buffer and a number of bytes to
receive.
Now, the application (ln~29--35) invokes the abstraction (ln~33) passing a
local buffer (ln~32).
%
The driver (ln~5--27) now hides the low level interface of the previous example
(ln~7--8) and implements the code abstraction (ln~15--27) that protects the
accesses to the shared buffer with an \code{atomic} block (ln~18--21).
%
The abstraction copies the driver buffer into the application buffer (ln~19)
up to the requested number of bytes (ln~22).
%
On the one hand the extra copying incurs a memory and runtime overhead, but on
the other hand it prevents race conditions with the ISR.

@@

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
This work was supported by the Serrapilheira Institute (grant number
Serra-1708-15612).
\end{acks}

>>>>

The addition of asynchronous ISRs now poses real threats concerning race
conditions since they interrupt execution at arbitrary points and could
possibly share memory with synchronous code.
%
Listing~\ref{lst.usart} illustrates a minimum serial communication (USART)
application (ln~15--24) that consumes incoming bytes (ln~19--23) as they arrive
(ln~18).
%
The USART API (ln~1--4) exposes an input event to signal incoming data (ln~3)
and a buffer to prevent data loss (ln~4).
The buffer is indispensable because the microcontroller might not cope with the USART 
speed.
%
The (unsafe) driver ISR (ln~6--13) simply appends incoming bytes to the end of the
buffer (ln~9) and signals the application (ln~11).
%
As a possible race condition, the ISR might fire and append a new byte to the
buffer (ln~9) just before the application clears it (ln~23), in which case the
new byte would be lost forever.

\CEU treats a \code{spawn~async/isr} as a block that runs in parallel with the
rest of the program (hence the prefix \code{spawn}).
%
This way, it is clear for the compiler that the accesses to \code{rx\_buf} in
Listing~\ref{lst.usart} may lead to race conditions (e.g., ln~9,23), and \CEU
raises a compile-time error.
%
The programmer is responsible to protect concurrent memory accesses with
\code{atomic} blocks, which disables interrupts for a short period
of time.

@@

However, we do not expect that application programmers should be required to
resolve race conditions.
\CEU provides \code{code} reactive abstractions (similar to coroutines) that help hiding
concurrency issues inside drivers.
%
Listing~\ref{lst.usart.code} changes the USART API (ln 1--3) to expose a code
abstraction that expects a reference to a buffer and a number of bytes to
receive.
Now the application (ln~29--35) invokes the abstraction (ln~33) passing a local
buffer (ln~32).
%
The driver (ln~5--27) now hides the low-level interface (ln~7--8) and
implements the abstraction (ln~15--27) that protects the accesses to the shared
buffer with an \code{atomic} block (ln~18--21).
%
The code copies the driver buffer into the application buffer (ln~19) up to the
requested number of bytes (ln~22).
%
On the one hand the extra copying incurs a memory and runtime overhead, but on
the other hand it prevents race conditions with the ISR.

@@

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}
